---
title: "12_bioconductor_single_cell"
author: "Jianwei cui"
date: '2022-05-25'
output: html_document
---

```{r, message=FALSE}
library( DropletUtils  )
library(tidyverse)
```

# Exercise

## Import scRNA-seq data and create a SingleCellExperiment object

- Import the filtered matrix into R; use `DropletUtils`.

**Note:** use the `samples=` argument of the `DropletUtils::read10xCounts()` function to give a memorable name to each sample.
  Check the difference without using the `samples` argument.

```{r}
library(DropletUtils)
sce <- DropletUtils::read10xCounts(samples =c(PBMC5K= "/project/obds/shared/resources/4_r_single_cell/pbmc_5k/filtered_feature_bc_matrix/"))
#sample.names argument can also give names to the sample files
#sample= are a directory contains three file barcodes features and matrix
```

- Print the object.
  What can you tell about its contents?
  
```{r}
sce
```

> Answer:
>gene ID is stoled and symbol
  
- What can you tell from the object metadata?

**Note:** slots of `SummarizedExperiment` objects are typically accessed using functions of the same name, e.g. `metadata()`.

```{r}
metadata(sce)
```

> Answer:
>

# Exercise

## Quality control

- Compute and visualise quality control metrics (library size, genes detected, mitochondrial fraction); use `scuttle` and/or `scater`.

  + Identify mitochondrial genes and pass those to the `subsets` argument of the `scuttle::addPerCellQC()` function.

  + What is the return value?
    Where are the quality metrics stored?
    What is the difference with `scuttle::perCellQCMetrics()`?

```{r}
is.mito<-grep(pattern="^MT-",rowData(sce)$Symbol)
 length(is.mito)
  is.mito
 #position is safer for the QC

```

```{r}
library(scuttle)     

sce <- scuttle::addPerCellQC( x= sce, subsets= list(percent_mt=is.mito))
colData(sce)
```

> Answer:
>

- Visualise library size, genes detected and mitochondrial fraction as three violin plots; use `ggplot2`.

```{r}
plot1 <- colData(sce) %>%
    as_tibble() %>% 
    ggplot() +
    geom_violin(aes( x=Sample,y=sum )) +
    labs(x = "Total UMI", y = "Value")
plot2 <- colData(sce) %>%
    as_tibble() %>% 
    ggplot() +
    geom_violin(aes(  x=Sample,y=detected   )) +
    labs(x = "Genes detected", y = "Value")
plot3 <- colData(sce) %>%
    as_tibble() %>% 
    ggplot() +
    geom_violin(aes( x=Sample,y=subsets_percent_mt_percent   )) +
    labs(x = "Percentage mitochondrial", y = "Value")
cowplot::plot_grid(plot1, plot2, plot3, nrow = 1)
```

- Filter cells, keeping those with more than 4,500 UMI, less than 15% mitochondrial UMI, and more than 1,500 genes detected. 

```{r}
table(sce$sum>4500&
sce$detected>1500&
sce$subsets_percent_mt_percent<15)
sce <- sce[   ,sce$sum>4500&
sce$detected>1500&
sce$subsets_percent_mt_percent<15]
sce
```

- Similarly, use `scuttle::perFeatureQCMetrics()` or `scuttle::addPerFeatureQC()` to compute per-feature quality metrics, and visualise those metrics.

```{r}
sce <- scuttle::addPerFeatureQC(sce   )
```

```{r}
## ggplot2
sce
rowData(sce)

ggplot(as.data.frame(rowData(sce)))+geom_point(aes(y=mean,x=detected))

```

# Exercise step 3. Normalisation

- Convert the counts into normalized expression values to eliminate cell-specific biases (e.g., in capture efficiency); use `scuttle` and/or `scran`.
  Display the names of the assays available after that step.

**Note:** use `scuttle::logNormCounts()` to compute log-normalised counts.
  What is the return value?
  Where can you find the normalised counts?

```{r}
library(scuttle)
sce <- scuttle::logNormCounts(sce   )
assayNames(sce)
```

> Answer:
> 

- Plot the variance against the mean of each gene.

**Note:** how can you tell whether the normalisation was effective?
  Compare with https://osca.bioconductor.org/feature-selection.html#quantifying-per-gene-variation

```{r}
library(DelayedMatrixStats)
#calculate row means and rowavrs in summarizedexpeiment data
x <- DelayedArray(assay(sce, "counts"))
plot_data <- tibble(
    mean = DelayedMatrixStats::rowMeans2(x),
    variance = DelayedMatrixStats::rowVars(x)
)
plot_counts <- ggplot( plot_data  ) +
    geom_point(aes(x=mean,y=variance))
#
x <- DelayedArray(assay(sce, "logcounts"))
plot_data <- tibble(
    mean = DelayedMatrixStats::rowMeans2(x),
    variance = DelayedMatrixStats::rowVars(x)
)
plot_logcounts <- ggplot(plot_data   ) +
    geom_point(aes(x=mean,y=variance))
cowplot::plot_grid(plot_counts, plot_logcounts, nrow = 1)
```

> Answer:
> 

- When would you rather use `scuttle::computePooledFactors` instead?

> Answer:
> 
> 

# Exercise

## Feature selection

Select features for downstream analyses, e.g. highly variable genes; use `scran`.

- Use `scran::modelGeneVar()` to model the variance of the log-expression profiles for each gene.
  What is the output?

```{r}
library(scran)
dec <- scran::modelGeneVar( sce  )
dec
```

> Answer:
> the output is a dataframe ,contains for the total technical and biological varaiance 

- Visualise the relation between the mean expression of each gene and the total / biological / technical variance of each gene.

How do you interpret those different values?

```{r}
ggplot(as_tibble(dec)) +
    geom_point(aes(mean, total), color = "black") +
    geom_point(aes(mean, bio), color = "blue") +
    geom_point(aes(mean, tech), color = "red")
#tech variance is the averange trend though the means increased 
```

> Answer:
> 

- Use `scran::getTopHVGs()` to identify highly variable genes (e.g., top 10%).

What is the output?
How many genes do you identify?
Where are those genes located in the mean vs. (biological) variance plot?

```{r}
hvg <- scran::getTopHVGs( dec ,
                          var.field = "bio",
                          n= 2000,
                          row.names=TRUE)
length(hvg)
head(hvg)
```


```{r}
## ggplot2
rownames(dec)
#add column in dec table for variable genes
dec$hvg<-rownames(dec)%in% hvg 
dec

ggplot(as_tibble(dec)) +
    geom_point(aes(mean, total), color = "black") +
    geom_point(aes(mean, bio), color = "blue") +
    geom_point(aes(mean, tech), color = "red") +
   geom_point(aes(mean, bio,color = hvg) ) 

#plot for variance plots


```

> Answer:
> 2000 genes and biological varaiance is more than 0
> 

# Exercise

## Dimensionality reduction

- Apply PCA; use `scater` or `BiocSingular`.
  Set a seed to control reproducibility.
  List the names of dimensionality reduction results available.

**Note:** only give the set of highly variable genes to the `scater::runPCA()` function, to save time, memory, and to focus on biologically informative genes in the data set.

```{r}
set.seed(1234)
sce <- scater::runPCA( sce,ncomponents=29, subset_row= hvg )
reducedDimNames(sce)
```

- Apply UMAP and t-SNE successively on the output of the PCA.
  List the names of dimensionality reduction results available each time.

**Note:** UMAP and t-SNE are typically given the output of PCA as their own input, to further reduce the dimensionality for ease of visualisation.

```{r}
sce <- scater::runUMAP( sce,dimred="PCA"  ,n_dimred=20)

```

```{r}
sce <- scater::runTSNE(  sce,dimred="PCA"  ,n_dimred=20  )
reducedDimNames(sce)
```

- Visualise the scatterplot of cells produced by each of those dimensionality reduction methods.
  Considering coloring points with quality control metrics.
  
```{r}





sce$log_sum <- log(sce$sum+1)
#reorder the column order randomlly 
sce <- sce[, sample(ncol(sce))]
sce_umap <- scater::plotReducedDim(sce,dimred="UMAP",colour_by = "log_sum")

sce_tsne<-scater::plotReducedDim(sce,dimred = "TSNE")




sce_umap
sce_tsne
```
  
## Bonus point

- Use `scran::denoisePCA()` to remove principal components that correspond to technical noise, and compare downstream t-SNE or UMAP with those obtained before de-noising.
  Name the output `sce_denoise`.
  How many components remain after denoising?
  Visualise a UMAP of the denoised PCA and compare.

```{r}
sce_denoise <- scran::denoisePCA(sce, assay.type = "logcounts" ,technical= dec,subset.row=hvg)

#overwhite the PCA components in the object
sce_denoise
dim(reducedDim(sce_denoise))
```

> Answer:
> denoise can remove the noise while it is not always the best approach to deal with the noise

```{r}
sce_denoise <- scater::runUMAP(sce_denoise,dimred="PCA"  ,n_dimred=5 )

```

```{r}
sce_denoise_umap <- scater::plotReducedDim(sce_denoise,dimred="UMAP",colour_by = "log_sum")

colnames(reducedDim(sce_denoise,"PCA"))


library(cowplot)

plot_grid(
    sce_umap + theme(legend.position = "bottom"),
    sce_denoise_umap + theme(legend.position = "bottom"),
    nrow = 1)
```

# Exercise

## Clustering

Cluster cells using `scran`.

- Start with `scran::getClusteredPCs()` to cluster cells after using varying number of PCs, and pick the number of PCs using a heuristic based on the number of clusters.

```{r}
output <- scran::getClusteredPCs(reducedDim(sce, "PCA"),max.rank = 40,by=2)
metadata(output)$chosen
output[8,]
```

- Use `scran::buildSNNGraph()` and `igraph::cluster_louvain()` with that "ideal" number of PCs.
  Assign the cluster label to a cell metadata column named `"label"`.

```{r, message=FALSE}
set.seed(1234)
sce <- scater::runPCA( sce,ncomponents=29, subset_row= hvg )
reducedDimNames(sce)

reducedDimNames(sce)
dim(reducedDim(sce,"PCA"))
#only run 29 on raw data for PCA analysis ,then can put the whole object into this buildSNNGraph function
g <-    scran::buildSNNGraph(sce,use.dimred="PCA")
g
colData(sce)[["label"]] <- igraph::cluster_louvain(g)$membership %>% as.factor()
  
  str(igraph::cluster_louvain(g))
```

- Visualise the assigned cluster on your preferred dimensionality reduction layout.

**Note:** Dimensionality reduction and clustering are two separate methods both based on the PCA coordinates.
  They may not always agree with each other, often helping to diagnose over- or under-clustering, as well as parameterisation of dimensionality reduction methods.

```{r}
gg_snn <- reducedDim(x = sce, type = "UMAP") %>%
    as.data.frame() %>%
    as_tibble() %>%
    bind_cols(colData(sce) %>% as_tibble()) %>%
    sample_frac() %>%
    ggplot() +
    geom_point(aes(V1, V2, color=label)) +
    cowplot::theme_cowplot()
gg_snn
```

## Bonus point

- Test different numbers of principal components and compare results.

```{r, message=FALSE}
snn_plots <- list()
for (d in c(5, 10, 13, 15)) {
    g <- scran::buildSNNGraph(t(reducedDim(sce, "PCA")), d = d)
    colData(sce)[[sprintf("snn_d", d)]] <- factor(igraph::cluster_louvain(g)$membership)
    gg_d <- reducedDim(x = sce, type = "UMAP") %>%
        as.data.frame() %>%
        as_tibble() %>%
        bind_cols(colData(sce) %>% as_tibble()) %>%
        sample_frac() %>%
        ggplot() +
        geom_point(aes(V1, V2, color=snn_d)) +
        labs(title = d) +
        cowplot::theme_cowplot()
    snn_plots[[as.character(d)]] <- gg_d
}
plot_grid(plotlist = snn_plots, ncol = 2)
```

- Try `scran::quickCluster()`; identify key parameters and compare results.

```{r}
sce$quickCluster <- scran::quickCluster(   )

gg_cluster <- reducedDim(x = sce, type = "UMAP") %>%
    as.data.frame() %>%
    as_tibble() %>%
    bind_cols(colData(sce) %>% as_tibble()) %>%
    sample_frac() %>%
    ggplot() +
    geom_point(aes(V1, V2, color=quickCluster)) +
    cowplot::theme_cowplot()
gg_cluster
```

# Exercise

## Cluster markers

- Use `scran::findMarkers()` to identify markers for each cluster.
  Display the metadata of markers for the first cluster.

```{r}
markers <- scran::findMarkers( sce,groups=colData(sce)$label  )
#all the markers finds contains both the positive FC and negative FC


####markers[[1] %>%lapply() %>%  arrange(FDR,desc(p.value))
        
        
        
colData(sce)
markers[[1]]
str(markers)
```

- Visualise the expression of selected markers:

  + As a dot plot, optionally with a violin layer.

```{r}
marker_id <-rownames(markers[[1]]) %>% head(n=10)
marker_name <-  rowData(sce)["ENSG00000163131","Symbol"]  

marker_name

library(scater)
plotDots(sce,features=marker_id,group="label")

plotExpression(sce,features=marker_id,x="label",colour_by = "label")

```

  + On a dimensionality reduction layout.
    Compare with the cluster labels.

```{r}
gg_marker <-  








plot_grid(gg_marker, gg_snn)
```

# Exercise

## Interactive visualisation

- Use `iSEE::iSEE()` to launch an interactive web-application to visualise the contents of the `SingleCellExperiment` object.

```{r}
library(iSEE)
app <- iSEE(sce)
if (interactive()) {
  shiny::runApp(app)
}
```

## Bonus point

- Preconfigure the application to start with a subset of panels, e.g.

```{r}
initial_panel_list <- list(
  ReducedDimensionPlot(PanelWidth=4L),
  RowDataTable(PanelWidth=8L)
)
app <- iSEE::iSEE(sce, initial = initial_panel_list)
if (interactive()) {
  shiny::runApp(app)
}
```