---
title: "13_singlecell_droplets"
author: "Jianwei cui"
date: '2022-05-26'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DropletUtils)
library(tidyverse)
library(ggplot2)
library(cowplot)
library(scDblFinder)
library(Seurat)
library(DoubletFinder)
library(celda)
library(scater)
```

# Exercise

## Empty droplets

- Import the raw matrix into R; use `DropletUtils`.
  Make a `SingleCellExperiment` object called `sce`.

```{r}
library(DropletUtils)
sce <- DropletUtils::read10xCounts(samples = c(
    "raw" = "/project/obds/shared/resources/4_r_single_cell/pbmc_5k/raw_feature_bc_matrix/"
))
  
  

sce
rowData(sce)
```

- Remove droplets without any count at all.

This will massively speed up downstream steps, while removing unquestionable empty droplets.

```{r}
library(tidyverse)
colSums(counts(sce)) %>% head()
which(colSums(counts(sce)) >0) %>% head()
#colSums(counts(sce)) >0 also works, return TRUe and FALSE
sce <- sce[,which(colSums(counts(sce)) >0) ]
sce
```
```{r}

rank_df<-DropletUtils::barcodeRanks(sce)
head(rank_df)

ggplot(as.data.frame(rank_df),aes(x=rank, y=total))+geom_point()+scale_x_log10()+scale_y_log10()+geom_hline(yintercept = 100,colour="red")
```

- Run `DropletUtils::emptyDrops()`.

```{r}
set.seed(123)
out <- DropletUtils::emptyDrops(sce)
out
#the everything is less than 100UMI, the columns will be NA have not been tested 
summary(out$Total>100)
subset(out,Total>100) %>% head()
ggplot(as.data.frame(out),aes(x=FDR))+geom_histogram()

```

**How do you read the output? Which droplets would you consider empty?**

> Answer:
> 
>
>
> 

- Remove empty droplets from the data sets.
  Call the new object `sce_filtered_emptydrops`.

```{r}
out$empty<-out$FDR<0.001
head(out)
which(out$empty==TRUE)
#return the location of the cell have FDR < 0.001

sce_filtered_emptydrops <- sce[,which(out$empty==TRUE)]
sce_filtered_emptydrops
```

## Bonus point

- Draw plots comparing empty droplets and other droplets
  (e.g., library size, mitochondrial content, PCA).

```{r}
is.mito <- grepl("^MT-", rowData(sce)$Symbol)
library(scuttle)
sce <- scuttle::addPerCellQC(sce, percent_top = 50, subsets = list(MT=is.mito))

colData(sce)[["empty"]] <-out$empty
#add a empty column true for cells, false for empty droplets and NA for empty dropet
plot1 <- colData(sce) %>%
    as_tibble() %>% 
    ggplot() +
    geom_violin(aes(empty, sum),scale= "width") +
    labs(x = "Is cell containing droplet", y = "total UMI")
plot2 <- colData(sce) %>%
    as_tibble() %>% 
    ggplot() +
    geom_violin(aes(empty, detected),scale= "width") +
    labs(x = "Is cell containing droplet", y = "features")
plot3 <- colData(sce) %>%
    as_tibble() %>% 
    ggplot() +
    geom_violin(aes(empty, subsets_MT_percent),scale= "width") +
    labs(x = "Is cell containing droplet", y = "MT percentage")
cowplot::plot_grid(plot1, plot2, plot3, nrow = 1)
```


```{r}
plot_data <- tibble(
  
  
)
ggplot(   ) +
  
  
  
```

```{r}
is.mt <- 
# table(is.mt)
plot_data <- tibble(
  
  
  
  
)
ggplot(   ) +
  
  
  
```

```{r}

sce <- 
sce <- 
plot_data <- tibble(
  
  
)
ggplot(   ) +
  
  
```

- Import the filtered matrix for the same cells and compare their set of filtered cells to yours.

```{r}
sce_filtered_cellranger <- 
  
  


all_barcodes <- colnames(sce)

compare_table <- tibble(
  emptydrops = all_barcodes %in%    ,
  cellranger = all_barcodes %in%    ,
)
table(compare_table)
```

# Exercise

## Doublets

Continue using the data set after you excluded empty droplets.

- Run `scDblFinder::scDblFinder()`.

```{r}
set.seed(123)
sce_scdblfinder <- scDblFinder::scDblFinder(sce=sce_filtered_emptydrops,clusters = TRUE)
#clusters=TRUE will run the cluster itself if no cluster has run
#clusters are used to defined that cell from different clusters are used to generate artifical doublets 
colData(sce_scdblfinder)
table(  colData(sce_scdblfinder)$scDblFinder.class )
plotDoubletMap(sce_scdblfinder)
```

**How do you read the output? Which droplets would you consider doublets?**

> Answer:
> 
> 
> 
> 

- Remove doublets from the data set.
  Call the new object `sce_filtered_scdblfinder`.

```{r}
sce_filtered_scdblfinder <- sce_scdblfinder[,colData(sce_scdblfinder)$scDblFinder.class=="singlet"]
sce_filtered_scdblfinder
```

## Bonus point

- Draw plots comparing doublets and other droplets.
  (e.g., library size, mitochondrial content, PCA)

```{r}



is.mito <- grepl("^MT-", rowData(sce_scdblfinder)$Symbol)
library(scuttle)
sce_scdblfinder <- scuttle::addPerCellQC(sce_scdblfinder, percent_top = 50, subsets = list(MT=is.mito))

colData(sce_scdblfinder)
#add a empty column true for cells, false for empty droplets and NA for empty dropet
plot1 <- colData(sce_scdblfinder) %>%
    as_tibble() %>% 
    ggplot() +
    geom_violin(aes(scDblFinder.class, sum),scale= "width") +
    labs(x = "cell class", y = "total UMI")
plot2 <- colData(sce_scdblfinder) %>%
    as_tibble() %>% 
    ggplot() +
    geom_violin(aes(scDblFinder.class, detected),scale= "width") +
    labs(x = "cell class", y = "features")
plot3 <- colData(sce_scdblfinder) %>%
    as_tibble() %>% 
    ggplot() +
    geom_violin(aes(scDblFinder.class, subsets_MT_percent),scale= "width") +
    labs(x = "cell class", y = "MT percentage")
cowplot::plot_grid(plot1, plot2, plot3, nrow = 1)

plot_data <- tibble(
  
  
)
ggplot(   ) +
  
  
  
```

```{r}
is.mt <- 

plot_data <- tibble(
  
  
  
  
)
ggplot(   ) +
  
  
  
```

```{r}
sce_scdblfinder <- 
sce_scdblfinder <- 
plot_data <- tibble(
  
  
)

ggplot(   ) +
  
  
```

- Compare results with the output of `chris-mcginnis-ucsf/DoubletFinder`

  + You will need to make a `Seurat` object.

```{r, message=FALSE}
## Pre-process Seurat object (standard)
seu <- 
seu <- 
seu <- 
seu <- 
seu <- 
seu <- 
```

```{r, message=FALSE}
## pK Identification (no ground-truth) 

sweep.res.list <- 

sweep.stats <- 
bcmvn <- 
```

```{r}
ggplot() +
  
  
  
```

```{r}
## Homotypic Doublet Proportion Estimate 
homotypic.prop <- 
nExp_poi <- 
nExp_poi.adj <- 
```

```{r, message=FALSE}
## Run DoubletFinder with varying classification stringencies 
seu <- 
```

```{r}
table(   )
```

```{r, message=FALSE}
seu <- 
```

```{r}
table(   )
```

```{r}
doublets_table <- data.frame(
  
  
  
)
table(doublets_table)
```

# Exercise

## Ambient RNA

- Run `celda::decontX()` (continue using the data after excluding empty droplets and doublets).

```{r, message=FALSE}
colData(sce_filtered_scdblfinder)
sce_decontX <- decontX(sce_filtered_scdblfinder,z=colData(sce_filtered_scdblfinder)$scDblFinder.cluster)
dim(sce_filtered_scdblfinder)
colnames(colData(sce_decontX))
ggplot(as.data.frame(colData(sce_decontX)),aes(x=decontX_contamination))+geom_histogram()
head(assay(sce_decontX,"decontXcounts"))
sce_decontX
```

- Visualize the UMAP layout created by `decontX`; use `plotDimReduceCluster()`.

```{r}
umap <- reducedDim(sce_decontX,"decontX_UMAP")
  head(umap)
plotDimReduceCluster(sce_decontX$decontX_clusters ,reducedDimName = "decontX_UMAP",dim=umap[,1],dim2 = umap[,2]  )     

```

```{r}
sce_decontX <- decontX(sce_filtered_scdblfinder,varGenes=2000)
#cannbe used in seurat object, need to put into intergral 
colData(sce_decontX)
ggplot(as.data.frame(colData(sce_decontX)),aes(x=decontX_contamination))+geom_histogram()
umap <- reducedDim(sce_decontX,"decontX_UMAP")
  head(umap)
plotDimReduceCluster(sce_decontX$decontX_clusters ,reducedDimName = "decontX_UMAP",dim=umap[,1],dim2 = umap[,2]  )   
plotDimReduceCluster(sce_decontX$decontX_clusters ,reducedDimName = "decontX_UMAP",dim=umap[,1],dim2 = umap[,2]  )  
```


- Visualize the percentage of contamination estimated by `decontX` in each cell.
  Use the `plotDecontXContamination()` function.

```{r}
plotDecontXContamination( sce_decontX  )
```

**Does any cluster display higher (estimated) levels of ambient RNA?**

> Answer:
> 
> 

- Visualize the log-normalised expression of marker genes on a UMAP layout faceted by gene; use `plotDimReduceFeature()`.

**Note:** For instance, use `CD3D`, `CD3E`, `GNLY`, `LYZ`, `S100A8`, `S100A9`, `CD79A`, `CD79B`, and `MS4A1`.

<!--
Source: https://bioconductor.org/packages/devel/bioc/vignettes/celda/inst/doc/decontX.pdf
(Section 5.3)
We will use:
- CD3D and CD3E for T-cells,
- LYZ, S100A8, and S100A9 for monocytes,
- CD79A, CD79B, and MS4A1 for B-cells,
- GNLY for NK-cells,
- and PPBP for megakaryocytes
-->



```{r}

sce_decontX<-logNormCounts(sce_decontX,assay.type ="decontXcounts")
assays(sce_decontX)
#make uqique name with gene ID and symbols
rownames(sce_decontX) <- uniquifyFeatureNames(ID=rowData( sce_decontX)$ID,names = rowData( sce_decontX)$Symbol

)
assay(sce_decontX,"logcounts")
plotDimReduceFeature(assay(sce_decontX,"logcounts"),
                     features = c("CD3D", "CD3E", "GNLY", "LYZ", "S100A8", "S100A9", "CD79A", "CD79B", "MS4A1"),
               
                    dim1=umap[,1],dim2 = umap[,2]
                     
  
  
  
  
  
  
  
)
```

- Visualise the percentage of cells with detectable expression of marker genes before and after decontamination as a bar plot; use `plotDecontXMarkerPercentage()`.

```{r}

markers <- list(
  
  Tcell_Markers = c("CD3E", "CD3D"), 
  Bcell_Markers = c("CD79A", "CD79B", "MS4A1"), 
  Monocyte_Markers = c("S100A8", "S100A9", "LYZ"), 
  NKcell_Markers = "GNLY" 
  
  
)
cellTypeMappings <- list(
  T_cell=c(1,3),B_cell=c(4),Mono_cell=c(2,9),NK=c(6)
  )
plotDecontXMarkerPercentage(sce_decontX,
                            markers = markers,
                            groupClusters = cellTypeMappings,
                            assayName = c("counts", "decontXcounts")
  
  
  
  
)
```



- Visualize the log-normalised expression of marker genes before and after decontamination as a violin plot; use `plotDecontXMarkerExpression()`.

```{r}
sce_decontX
#similar plot just a violin plot
plotDecontXMarkerExpression(
  sce_decontX,
  markers = markers,
                            groupClusters = cellTypeMappings,
                            assayName = c("counts", "decontXcounts"),
  log1p=TRUE
  
  
  
  
)
```
