---
title: "ggplot2_exercise and object oritented programming"
author: "Jianwei cui"
date: '2022-05-12'
output: html_document
---
```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Demo

## Character

```{r}
character_object <- "A"
character_object
```

```{r}
str(character_object)
```

```{r}
class(character_object)
```

```{r}
attributes(character_object)
```

```{r}
character_object2 <- c("A", "B", "C")
character_object2
```

```{r}
str(character_object2)
```

```{r}
class(character_object2)
```

```{r}
attributes(character_object2)
```

# Demo

## Factors

```{r, include=TRUE}
factor_object <- factor(c("A", "B", "C"))
factor_object
```

```{r}
str(factor_object)
```

```{r}
class(factor_object)
```

```{r}
attributes(factor_object)
```

```{r}
attr(factor_object, "levels")
```

```{r}
attr(factor_object, "class")
```

# Demo

## Data-frames

```{r}
dataframe_object <- data.frame(
  numbers = c(1, 2, 3),
  strings = c("A", "B", "C")
  )
dataframe_object
```

```{r}
class(dataframe_object)
```

```{r, include=TRUE}
str(dataframe_object)
```

```{r}
attributes(dataframe_object)
```

# Demo

## Matrices

```{r}
matrix_object <- matrix(
  data = c(1, 2, 3, 4, 5, 6, 7, 8),
  nrow = 2
  )
matrix_object
```

```{r}
class(matrix_object)
```

```{r}
str(matrix_object)
```

```{r}
attributes(matrix_object)
```

```{r}
attr(matrix_object, "dim")
```

# Demo

## S3 method dispatch

```{r}
methods("print") %>% head(14)
```

```{r}
print.factor(factor_object)
```

```{r}
print.factor(character_object)
```

```{r}
print.default(character_object)
```


## ggplot2 exercise 
```{r}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggrepel)
library(cowplot)
library(patchwork)
library(gridExtra)
```

# Exercise

Use `library(ggplot2)` to load the package.

```{r}
library(ggplot2)
```

## Specify the dataset.

What does `ggplot(diamonds)` do?

```{r}
ggplot(diamonds)
```

> Answer:

## Add the aesthetics.

What does `ggplot(diamonds, aes(x = carat, y = price))` do?

```{r,fig.height=12,fig.width=12}
ggplot(diamonds, aes(x = carat, y = price))
```

> Answer:

## Add geometric objects

- Add data points showing `carat` on the x-axis and `price` on the y-axis.

```{r}
ggplot(diamonds, aes(x = carat, y = price))+geom_point()
  
```

- Color data points by `cut`.
  (Copy-paste and extend the code chunk above.)

```{r}
ggplot(diamonds, aes(x = carat, y = price, colour = cut))+geom_point()
  
```

- Add a smoothed mean trend line.
  (Copy-paste and extend the code chunk above.)

```{r}

ggplot(diamonds, aes(x = carat, y = price, colour = cut))+geom_point()+geom_smooth()

```

- Assign that last plot to an object called `obds_diamonds`.

```{r}
obds_diamonds <- ggplot(diamonds, aes(x = carat, y = price, colour = cut))+geom_point()+geom_smooth()


```


# Exercise

Predict the difference between these two plots

### Plot 1

```{r, include=TRUE}
ggplot(diamonds, aes(x = carat, y = price, colour = cut)) +
  geom_point() +
  geom_smooth()?
```

### Plot 2

```{r, include=TRUE}
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point(aes(colour = cut)) +
  geom_smooth()
```

> Answer:

# Exercise

## Trend lines

Using the `ChickWeight` data set:

- Create a scatter plot of weight (y-axis) over time (x-axis).

```{r}
ggplot(ChickWeight,aes(x= Time, y= weight))+ geom_point()

```

- Color by diet.
  (Copy-paste and extend the code chunk above.)

```{r}

ggplot(ChickWeight,aes(x= Time, y= weight, color= Diet))+ geom_point()
```

- Add a linear mean trend line for each diet.
  (Copy-paste and extend the code chunk above.)

```{r}
ggplot(ChickWeight,aes(x= Time, y= weight, color= Diet))+ geom_point()+geom_smooth(method=lm)


```

At this point you should be able to visually identify
the diet that leads to the highest mean increase in weight.

> Answer:

- Facet a ribbon of sub-plots, one per diet.
  (Copy-paste and extend the code chunk above.)

```{r}

ggplot(ChickWeight,aes(x= Time, y= weight, color= Diet))+ geom_point()+geom_smooth(method=lm)+facet_wrap(~Diet,ncol = 2)


```

- Assign that last plot to an object called `obds_chickweight`.

```{r}

obds_chickweight<- ggplot(ChickWeight,aes(x= Time, y= weight, color= Diet))+ geom_point()+geom_smooth(method=lm)+facet_wrap(~Diet,ncol=2)


```

# Exercise

## Bar plot

- Load the `ggplot2::msleep` data set.

```{r}
ggplot2::msleep

```

- Draw a bar plot of number of observations (i.e., rows) for each taxonomic order
  (i.e, one plot and one bar per taxonomic order).

```{r,fig.width=14}
ggplot(ggplot2::msleep,aes(x= order)) + geom_bar()

```

- Change the angle and font size of the text for the x-axis ticks (not the axis titles).
  Justify the text of those x-axis ticks as right-aligned.
  (Copy-paste and extend the code chunk above.)

```{r}
ggplot(ggplot2::msleep,aes(x= order)) + geom_bar()+theme(axis.ticks= element_line(size=4),axis.text.x = element_text(size=9,angle = 45,hjust = 1))




```

- Change the value and font size of the title for both x and y axes.
  (Copy-paste and extend the code chunk above.)

```{r}

ggplot(ggplot2::msleep,aes(x= order)) + geom_bar()+theme(text=element_text(size =14), axis.ticks= element_line(size=4),axis.text.x = element_text(size=9,angle = 45,hjust = 1))+labs(x="Taxanomic order",y="counts of animal",title = "msleep dataset")





```

- Fill each bar with colors, proportionally to the count of each genus.
  (Copy-paste and extend the code chunk above.)

From this point onwards, you may need to iteratively resize the text of the ticks and axes for readability.

```{r}
ggplot(ggplot2::msleep,aes(x= order,fill = genus)) + geom_bar()+theme(text=element_text(size =14), axis.ticks= element_line(size=4),axis.text.x = element_text(size=9,angle = 45,hjust = 1))+labs(x="Taxanomic order",y="counts of animal",title = "msleep dataset")






```

- Reduce the legend key size.
  (Recommendation: use `unit(2, "mm")`).
  (Copy-paste and extend the code chunk above.)

```{r}
ggplot(ggplot2::msleep,aes(x= order,fill = genus)) + geom_bar()+theme(text=element_text(size =14), axis.ticks= element_line(size=4),axis.text.x = element_text(size=9,angle = 45,hjust = 1), legend.key.size = unit(2,"mm") )+labs(x="Taxanomic order",y="counts of animal",title = "msleep dataset")


```

- Force the legend to be display in 3 columns.
  (Recommendation: use `guide_legend(...)`).
  (Copy-paste and extend the code chunk above.)

```{r}

ggplot(ggplot2::msleep,aes(x= order,fill = genus)) + geom_bar()+theme(text=element_text(size =14), axis.ticks= element_line(size=4),axis.text.x = element_text(size=9,angle = 45,hjust = 1), legend.key.size = unit(2,"mm") )+labs(x="Taxanomic order",y="counts of animal",title = "msleep dataset")+ guides(fill= guide_legend(ncol=3))








```

- Add a contour of thin black lines to the bars.

```{r}

ggplot(ggplot2::msleep,aes(x= order,fill = genus)) + geom_bar(colour= "black",size=0.1)+theme(text=element_text(size =14), axis.ticks= element_line(size=4),axis.text.x = element_text(size=9,angle = 45,hjust = 1), legend.key.size = unit(2,"mm") )+labs(x="Taxanomic order",y="counts of animal",title = "msleep dataset")+ guides(fill= guide_legend(ncol=3))



```

- Assign that last plot to an object called `obds_msleep`.

```{r}



obds_msleep <- ggplot(ggplot2::msleep,aes(x= order,fill = genus)) + geom_bar(colour= "black",size=0.1)+theme(text=element_text(size =14), axis.ticks= element_line(size=4),axis.text.x = element_text(size=9,angle = 45,hjust = 1), legend.key.size = unit(2,"mm") )+labs(x="Taxanomic order",y="counts of animal",title = "msleep dataset")+ guides(fill= guide_legend(ncol=3))



```

# Exercise

## Plotting grid

- Collate the plots that we assigend to objects through the day, as a single plot.

  + Plots: `obds_diamonds`, `obds_chickweight`, `obds_msleep`.
  
  + Methods: `cowplot::plot_grid()`, `patchwork`, `gridExtra::grid.arrange()`.

Using `cowplot`.

```{r}
library(cowplot)
first_row <- cowplot::plot_grid( obds_diamonds,obds_chickweight,labels = c("A","B"),ncol = 2
)

second_row <- cowplot::plot_grid(obds_msleep,ncol = 1,labels = c("C")
  
)
second_row
super_plot <- cowplot::plot_grid(first_row,second_row,nrow = 2
  
  
)
super_plot
```

Using `patchwork`.

```{r, fig.height=20, fig.width=12}
library(patchwork)
obds_diamonds + obds_chickweight
pach_plot<-((obds_diamonds + obds_chickweight)/obds_msleep)
```

Using `gridExtra`.

```{r, fig.height=12, fig.width=12}
library(gridExtra)
gridExtra::grid.arrange( obds_diamonds, obds_chickweight, ncol=2, nrow =1)


layout_matr<-rbind(c(1,3),c(2,3))
gridExtra::grid.arrange( obds_diamonds, obds_chickweight,obds_msleep,layout_matrix = layout_matr)
```

- Export the new plot in a PDF file, and open it in a PDF viewer (e.g. Adobe Acrobat Reader DC).

You will likely need a few attempts to fine-tune the width and height of the output file.

```{r}
ggsave(filename = "ggplot_exercise_plots.pdf",plot = pach_plot, width = 20,height = 20  )
ggsave(filename = "ggplot_exercise_plots_cowplot.pdf",plot = super_plot, width = 20,height = 20  )
```

# Exercise

## Pair programming

- Explore the data set `ggplot2::mpg` and generate the most informative plot that you can!

```{r}
ggplot2::mpg

ggplot(mpg,aes(x= cty, y= hwy,color=year ))+ geom_point()+geom_abline(intercept = 0,slope = 1,size=0.5)+theme(panel.grid.minor = element_blank(), text=element_text(size =14), axis.ticks= element_line(size=4),axis.text.x = element_text(size=9,angle = 45,hjust = 1),plot.title = element_text(hjust = 0.5))+facet_wrap(~manufacturer)+labs(y="highway usage",x="city usage",title = "Main title",subtitle = "subtitle")
head(mpg)
```
```{r}
df1 <- data.frame(col1 = c(1, 2, 3), col2 = c(6, 5, 4))
df2 <- data.frame(column1 = c(1.5, 1.75), column2 = c(3.5, 4))
df3 <- data.frame(variable1 = runif(100, min = 1, max = 3))
ggplot() +
geom_histogram(aes(x = variable1), df3, alpha = 0.5, binwidth = 0.5) +
geom_point(aes(x = col1, y = col2), df1, color = "blue") +
geom_point(aes(x = column1, y = column2), df2, color = "red") +
geom_label(aes(x = column1, y = column2, label = column1), df2, color = "red")
```

